Invasive lionfish present region-wide variation of allometric growth in the Western Atlantic

### 2018-04-12 docs/Manuscript.tex
Incorporated Sean's comments

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
North-Western Atlantic and the Caribbean. In order to better manage the
invasion, we must be able to accurately estimate their total biomass.
This work compares length-weight relationships of the invasive lionfish
through the invasion range. A review of 15 length-weight relationships
reported in 11 peer-reviewed studies shows that lionfish exhibit spatial
variation in weight-at-length. The reviewed parameters indicate that,
for the same length, lionfish in the Caribbean have lower body mass than
in the Atlantic or Gulf of Mexico. This highlights the importance of
using site-specific parameters to estimate biomass from length
observations. This study also reports a new pair of length-weight
parameters (\(a = 3.2056\times 10^{-6}; b = 3.235\)) for organisms
sampled in the Central Mexican Caribbean region. These fidnings can have
major implications in management, especially when estimating biomass
available for harvest or predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs.
\end{abstract}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

@@

At least 84\% of the marine eco-regions have reported the presence of an
invasive species \citep{molnar_2008}, which represent a major threat to
local biodiversity and the economic activities that depend on it
\citep{bax_2003}. Invasive species may threaten native species through
predation, competition, or indirect habitat effects
\citep{davis_2003, gurevitch_2004}. By 2005, the economic cost of
invasive species to the United States was estimated at USD\$120 billion
per year\citep{pimentel_2005}.

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the Western Atlantic and the Caribbean, likely introduced through
liberation of aquarium-kept organisms \citep{betancurr_2011}. They are
the first marine vertebrates to establish in North Atlantic
\citep{schofield_2009,schofield_2010} and Caribbean coasts
\citep{sabidoitza_2016}. Lionfish have been widely reported in coral
reefs \citep{aguilarperera_2010}, but also in other habitats such as
estuaries \citep{jud_2011}, mangroves \citep{barbour_2010}, areas with
hard-bottoms \citep{muoz_2011}, and mesophotic reefs
\citep{andradibrown_2017}. Due to their threat to local biodiversity,
the speed of their spread, and difficulty of management, their presence
in these waters has been labeled as a major marine invasion
\citep{hixon_2016}.

A substantial amount of research has been done to describe lionfish
feeding ecology in North Carolina \citep{muoz_2011}, the Bahamas
\citep{morris_2009,cote_2013}, Northern Gulf of Mexico
\citep{dahl_2014}, Mexican Caribbean
\citep{valdezmoreno_2012,villaseorderbez_2014}, Belize
\citep{hackerott_2017}, and Costa Rica \citep{sandel_2015}. Their
feeding behavior and high consumption rates can reduce recruitment
\citep{albins_2008} and population sizes \citep{green_2012} of native
reef-fish species, and further the endangerment of critically endangered
reef fish \citep{rocha_2015}. (However, see \citet{hackerott_2017} for a
case where there was no evidence that lionfish affected the density,
richness, or community composition of prey fishes). Major efforts have
been made to understand the possible impacts of the invasion by keeping
track of its range through time \citep{schofield_2009,schofield_2010}
and predicting invasion ranges under future climates
\citep{grieve_2016}. By combining information from these disciplines,
researchers have been able to predict the trophic impacts of lionfish
\citep{ariasgonzalez_2011}, which can then be translated into
ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit
organizations have promoted removal programs and incentivized its
consumption \citep{chin_2016}. In some cases, these have shown to
significantly reduce --but not quite eliminate-- lionfish abundances at
local scales \citep{sandel_2015,chin_2016,deleon_2013}. The rapid
recovery rates exhibited by lionfish \citep{barbour_2011} and the
persistent populations in mesophotic coral ecosystems
\citep{andradibrown_2017} --which can contribute with recruitment to
shallow-water populations-- make of complete eradication through fishing
effort an unlikely solution. However, further incentivizing its
consumption might create a demand big enough to promote and sustain a
stable fishery \citep{chin_2016}, which can reduce local abundances and
control the invasion while providing alternative livelihoods.

The feasibility of lionfish removal programs has been extensively
evaluated through field
observations\citep{usseglio_2017,sandel_2015,chin_2016,deleon_2013} and
empirical modeling \citep{barbour_2011,morris_2011,johnston_2015}. The
latter approeach models changes in biomass in response to changes in
mortality (\emph{i.e.} culling). In this case, biomass represents the
sum of all fish's individual weight. The individual weight of an
organism (Total Weight; TW) can be estimated from its Total Length (TL)
using the allometric growth equation (\(TW = aTL^b\)). Parameters \(a\)
and \(b\) for this equation exist for North Carolina
\citep{barbour_2011}, Northern \citep{fogg_2013,dahl_2014} and Southern
Gulf of Mexico \citep{aguilarperera_2016}, the Southern Mexican
Caribbean \citep{sabidoitza_2016}, Bahamas \citep{darling_2011}, Little
Cayman \citep{edwards_2014}, Jamaica \citep{chin_2016}, Bonaire
\citep{deleon_2013} Puerto Rico \citep{toledohernndez_2014}, and Costa
Rica \citep{sandel_2015}, but remain unavailable for the central Mexican
Caribbean. The length-weight of a species can vary across regions as a
response to biotic (\emph{e.g.} local food availability) and abiotic
(\emph{e.g.} water temperature) conditions \citep{johnson_2016}. Thus,
when using biomass-informed models or estimating weights from length
observations, it is important to use site-specific parameters. This is
especially important when research involves identifying the total
biomass available for harvest by fishers \citep{chin_2016} or the
efficacy of lionfish removals
\citep{barbour_2011,morris_2011,johnston_2015}.

This study reviews lionfish allometric growth throughout the invasion
range, and provides a new pair of parameters specific to the central
Meixcan Caribbean. The results suggest there are important
regional-scale variations in allometric growth patterns of lionfish. The
observed differences highlight the importance of using site-specific
parameters, especially when informing invasion management strategies.

@@

The main objective of this work was to compare allometric growth of
lionfish throughout the invasion range. Allometric parameters were
retreived from scientific literature, and an additional pair of
parameters was calculated from filed observations in the central Mexican
Caribbean.

Length-weight relationships (n = 15) identified in literature were
obtained for the North Atlantic (n = 1, \citet{barbour_2011}), Gulf of
Mexico (n = 7, \citet{aguilarperera_2016,fogg_2013,dahl_2014}), and
Caribbean (n = 8,
\citet{chin_2016,toledohernndez_2014,deleon_2013,edwards_2014,sabidoitza_2016,sandel_2015,darling_2011}).
When available, information on sampling methods, sex differentiation,
location, and depth ranges of each study was retrieved. Whenever gender
was not specified, it was assumed that the results were presented for
pooled genders. Locations with allometric parameters are shown in Figure
\ref{fig:map} and \ref{tab:all_params}.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Locations where
allometric growth of lionfish (\emph{Pterois spp}) have been reported.
Sizes indicate sample size of each study, colors indicate the b
coefficient in Eq. \ref{eq:allometric}.}
\end{figure}

Parameters from the central Mexican Caribbean were obtained from data
collected in 10 sampling sites along the Central Mexican Caribbean coast
in 2010 (Table S1). Sampling locations included wall and carpet reefs at
depths between 5.7 m and 38.1 m. All observed lionfish (n = 109) were
collected using hand nets and numbered collection bottles. The use of
hand nets prevented any weight loss due to bleeding and allowed a better
representation of small sizes by eliminating gear selectivity. Organisms
were euthanized (via pithing) and Total Length (TL; mm) and Total Weight
(TW; gr) were recorded before freezing organisms.

@@

Where \(TW\) is the Total Weight (gr), \(TL\) is the Total Length (mm),
\(a\) is the ponderal index and \(b\) is the scaling exponent or
allometric parameter. When \(b = 3\), it is said that the organism
exhibits a perfect isometric growth. The dependent and independent
variables were transformed via base-10 logarithms, thus the equation
becomes:

\begin{equation}
\label{eq:log-alo}
log_{10}(TW) = b\times log_{10}(TL) + log{10}(a)
\end{equation}

@@

Where \(Y = log_{10}(TW)\), \(X = log_{10}(TL)\), \(m = b\), and
\(c = log_{10}(a)\). Since \(b = m\), we will only use \(b\) throughout
the paper for simplicity. The coefficients (\(c\) and \(b\)) were
estimated with an Ordinary Least Square Regression and
heteroskedastic-robust standard error correction \citep{zeileis_2004}.
Both coefficients were tested against the null hypothesis of no change
(\emph{i.e.} \(H_0: c = 0\) and \(H_0: b = 0\)) and the null hypothesis
of isometric growth (\(H_0: b = 3\)). Coefficients were tested with a
two-tailed Student's t-test. The significance of the regression was
corroborated with an F-test.

During the review process, some studies indistinctly used \(a\) to
report either the ponderal index in \ref{eq:allometric} or the
y-intercept (\(c\)) in \ref{eq:log-alo-trans}. Others reported their
parameters as mm-to-gr conversions, but a rapid evaluation of such
parameters indicated that they were estimated as cm-to-gr conversions.
Here, all parameters are reported as TL(mm) to TW(gr) conversions. When
required, values from other studies were transformed for consistency.

Since uncertainty around estimated relationships was not reported in
some of the reviewed studies, it was not possible to test for
statistical differences between relationships. Instead, the 16
length-weight relationships were used to calculate predicted weight from
length observations of the organisms sampled from the Central Mexican
Caribbean. Predicted weights were divided by the observed weights to
obtain a ratio. Difference in mean weight ratios across studies were
tested with a one-way Analysis of Variance (ANOVA). All analyses were
performed in R version 3.4.4 \citep{rcore_2018}. Raw data and code used
in this work is available at dryad.org.

@@

This section presents a new length-weight relationship for lionfish in
the Central Mexican Caribbean region. A summary of length-weight
relationship is provided, and a comparison of predicted-to-observed
weight is included.

The model adjusted to Eq. \ref{eq:log-alo-trans} estimated the
coefficient values at \(b = 3.2347391\) and \(c = -5.4940866\). The
intercept (\(c\)) and slope \((b)\) were significantly different from
zero (\(t(107) = -66.17; p<0.001\) and \(t(107) = 83.24; p<0.001\),
respectively), rejecting the null hypothesis of no change. Additionally,
the allometric factor (\(b\)) was significantly different from the value
of isometric growth of \(b = 3\) (\(t(107) = 6.04; p<0.001\)),
indicating that lionfish present allometric growth. More information on
model fit is presented in TableS2. The relationship between Total Length
and Total Weight is presented in Figure \ref{fig:l-w-carib}.

From the 11 peer-reviewed studies including information on growth
parameters for \emph{P. volitans} and the additional calculated for the
central Mexican Caribbean, 15 parameters were identified (Table
\ref{tab:all_params}, Fig \ref{fig:all_allo}). Two studies
\citep{aguilarperera_2016,fogg_2013} reported gender-level and pooled
parameters, while the rest presented pooled results. The smallest
coefficient of determination was presented by \citet{chin_2016} with
\(R^2 = 0.8715\), while \citet{sabidoitza_2016} reported the highest
value at \(R^2 = 0.9907\). Reviewed studies presented information for
organisms obtained at depths between 0.5 and 57 m. Three studies
\citep{aguilarperera_2016,chin_2016,dahl_2014} explicitly stated that
their organisms were sampled with pole spears. Five studies
\citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,toledohernndez_2014}
mentioned that some of their organisms were obtained with pole spears
(or other type of harpoon), but also hand-held nets or fish traps. Two
studies \citep{deleon_2013,darling_2011} did not specify how organisms
were sampled. \citep{fogg_2013} stated the use spine-less weight in the
length-weight relationship estimation.

@@

There were significant differences in predicted-to-observed weight
ratios estimated for each pair of parameters (F(15, 1728) = 38.26; p
\textless{} 0.001). From all allometric parameters reviewed, those of
\citet{edwards_2014} provided the lowest weight estimates, with an
predicted-to-observed weight ratio of 0.98 \(\pm\) 0.23 (mean \(\pm\)
SD). On the other hand,\citet{barbour_2011} yielded the highest weight
estimates, with a mean (\(\pm\) SD) predicted-to-observed weight ratio
of 1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios and groups
identified by Tukey's HSD (\(\alpha = 0.05\)) are presented in Figure
\ref{fig:bio_ratio}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}\label{fig:l-w-carib}Length-weight
relationship for 109 lionfish sampled in the Central Mexican Caribbean.
Points indicate samples, solid line indicates curve of best fit,
marginal plots represent the density distribution of each variable.}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-4}\label{tab:all_params}Summary of 13 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies if data is for pooled (B), females (F), or males (M), a = scaling parameter for Eq. 1, c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The $R^2$ column indicates reported model fit.}
\centering
\begin{tabular}[t]{lllrrll}
\toprule
Region & Sex & n & a & b & R2 & Reference\\
\midrule
Caribbean & B & 458 & 3.64e-05 & 2.81 & - & Sandel et al., 2015\\
Caribbean & B & 419 & 2.75e-05 & 2.85 & 0.8715 & Chin et al 2016\\
Caribbean & B & 1450 & 2.28e-05 & 2.89 & 0.96 & de Leon et al., 2013\\
Caribbean & B & 1887 & 3.00e-06 & 3.24 & 0.97 & Edwards et al., 2014\\
Caribbean & B & - & 2.50e-06 & 3.29 & - & Darling et al., 2011\\
\addlinespace
Caribbean & B & 2143 & 5.20e-06 & 3.18 & 0.9907 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 8.00e-06 & 3.11 & 0.958 & Toledo-Hernández et al., 2014\\
Caribbean & B & 109 & 3.20e-06 & 3.23 & 0.9766 & This study\\
GoM & B & 934 & 2.10e-06 & 3.34 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 2.90e-06 & 3.30 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\addlinespace
GoM & F & 67 & 1.20e-06 & 3.47 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & M & 59 & 4.20e-06 & 3.23 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 1.40e-06 & 3.43 & 0.99 & Fogg et al., 2013\\
GoM & M & 119 & 2.70e-06 & 3.31 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 6.80e-06 & 3.14 & 0.94 & Fogg et al., 2013\\
NorthAtlantic & B & 774 & 2.89e-05 & 2.89 & - & Barbour et al 2011\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}\label{fig:all_allo}Length-weight
relationships (n = 16) for 11 studies and this study. Colors indicate
studies from which the parameters were extracted. Solid lines indicate
that the fit was performed for males and females pooled together. Dotted
lines indicate that the regression was performed on females, and dashed
lines indicate it was performed for males.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}\label{fig:bio_ratio}Violin plot showing
the distribution of predicted to observed weight ratios for 16 pairs of
allometric parameters. Red and blue circles indicate median and mean
values, respectively. Like letters indicate values that do not differ
significantly (Tukey's HSD; p \textless{} 0.05).}
\end{figure}

@@

A new pair of allometric growth parameters for lionfish in the central
Mexican Caribbean are provided. This compliments existing literature for
other sites in the north (\emph{i.e.} Alacranes Reef
\citep{aguilarperera_2016}) and South (Xcalak \citep{sabidoitza_2016})
of the Mexican Caribbean. Additionally, the study identifies regional
differences in length-weight relationships. Here, we focus the
discussion on plausible causes of these differences.

The length-weight coefficients estimated in this study were within the
range identified by studies in other regions \ref{tab:all_params}.
However, the ones presented here provide lower weight estimates for a
same length. Until about TL = 200 mm, there are no appreciable
differences between the parameters for organisms from the Mexican
Caribbean and those for little Cayman \citep{edwards_2014} and Jamaica
\citep{chin_2016}. Yet, for larger organisms (TL \textgreater{} 270 mm)
parameters from Costa Rica \citep{sandel_2015} and Bonaire
\citep{deleon_2013} provide similar estimates to those from this study.
Conversely, these same studies tend to estimate higher weights --as
compared to the ones reported here-- for smaller organisms, likely due
to the lack of small organisms in the samples used to estimate their
parameters.

>>>>

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
Western Atlantic and the Caribbean. In order to better manage the
invasion, we must be able to accurately estimate their total biomass.
This work compares length-weight relationships of the invasive lionfish
through the invasion range. A review of 15 length-weight relationships
reported in 11 peer-reviewed studies shows that lionfish exhibit spatial
variation in weight-at-length. The reviewed parameters indicate that,
for the same length, lionfish in the Caribbean have lower body mass than
in the Atlantic or Gulf of Mexico. This highlights the importance of
using site-specific parameters to estimate biomass from length
observations. This study also reports a new pair of length-weight
parameters (\(a = 3.2056\times 10^{-6}; b = 3.235\)) for organisms
sampled in the central Mexican Caribbean region. These findings can have
major implications in management, especially when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs.
\end{abstract}

\begin{document}

@@

At least 84\% of marine eco-regions have reported the presence of an
invasive species \citep{molnar_2008}, which represent a major threat to
local biodiversity and the economic activities that depend on it
\citep{bax_2003}. Invasive species may threaten native species through
predation, competition, or indirect habitat effects
\citep{davis_2003, gurevitch_2004}. By 2005, the economic cost of
invasive species to the United States was estimated at USD \$120 billion
per year \citep{pimentel_2005}.

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the western Atlantic and the Caribbean, likely introduced through
liberation of aquarium-kept organisms \citep{betancurr_2011}. They are
the first invasive marine vertebrates established along the North
Atlantic Caribbean coasts
\citep{schofield_2009,schofield_2010,sabidoitza_2016}. Lionfish have
been widely reported in coral reefs \citep{aguilarperera_2010}, but also
in other habitats such as estuaries \citep{jud_2011}, mangroves
\citep{barbour_2010}, hard-bottomed areas \citep{muoz_2011}, and
mesophotic reefs \citep{andradibrown_2017}. Their presence in these
waters has been labeled as a major marine invasion because they threaten
local biodiversity, spread rapidly, and are difficult to manage
\citep{hixon_2016}.

A substantial amount of research has been done to describe lionfish
feeding ecology in North Carolina \citep{muoz_2011}, the Bahamas
\citep{morris_2009,cote_2013}, Northern Gulf of Mexico
\citep{dahl_2014}, Mexican Caribbean
\citep{valdezmoreno_2012,villaseorderbez_2014}, Belize
\citep{hackerott_2017}, and Costa Rica \citep{sandel_2015}.
\citet{peake_2018} show that invasive lionfish prey on at least 167
different species across on the tropical and temperate North Atlantic.
Their feeding behavior and high consumption rates can reduce recruitment
and population sizes of native reef-fish species, and can further
endanger reef fish \citep{albins_2008, green_2012,rocha_2015}. (However,
see \citet{hackerott_2017} for a case where there was no evidence that
lionfish affected the density, richness, or community composition of
prey fishes). Major efforts have been made to understand the possible
impacts of the invasion by keeping track of its range through time
\citep{schofield_2009,schofield_2010} and predicting invasion ranges
under future climates \citep{grieve_2016}. By combining information from
these disciplines, researchers have been able to predict the trophic
impacts of lionfish \citep{ariasgonzalez_2011}, which can then be
translated into ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit
organizations have promoted removal programs and incentivized its
consumption \citep{chin_2016}. In some cases, these have shown to
significantly reduce --but not quite eliminate-- lionfish abundances at
local scales \citep{sandel_2015,chin_2016,deleon_2013}. Complete
eradication of lionfish through fishing is unlikely because of their
rapid recovery rates and ongoing recruitment to shallow-water areas from
their persistent populations in mesophotic coral ecosystems
\citep{barbour_2011,andradibrown_2017}. However, further incentivizing
its consumption might create a demand big enough to promote and sustain
a stable fishery \citep{chin_2016}, which can reduce local abundances
and control the invasion while providing alternative livelihoods.

The feasibility of lionfish removal programs has been extensively
evaluated through field observations
\citep{usseglio_2017,sandel_2015,chin_2016,deleon_2013} and empirical
modeling \citep{barbour_2011,morris_2011,johnston_2015}. The latter
approach models changes in biomass in response to changes in mortality
(\emph{i.e.} culling). In this case, biomass represents the sum of all
fish's individual weight. The individual weight of an organism can be
estimated from its length using the allometric growth equation.
Parameters for this equation exist for North Carolina, Northern and
Southern Gulf of Mexico, the Southern Mexican Caribbean, Bahamas, Little
Cayman, Jamaica, Bonaire, Puerto Rico, and Costa Rica, but remain
unavailable for the central Mexican Caribbean
\citep{barbour_2011,fogg_2013,dahl_2014,aguilarperera_2016,sabidoitza_2016,darling_2011,edwards_2014,chin_2016,deleon_2013,toledohernndez_2014,sandel_2015}.
The length-weight of a species can vary across regions as a response to
biotic (\emph{e.g.} local food availability) and abiotic (\emph{e.g.}
water temperature) conditions \citep{johnson_2016}. Thus, when using
biomass-informed models or estimating weights from length observations,
it is important to use site-specific parameters. This is especially
important when research involves identifying the total biomass available
for harvest by fishers or the efficacy of lionfish removals
\citep{barbour_2011,morris_2011,johnston_2015,chin_2016}.

This study reviews lionfish allometric growth throughout the invasion
range, and provides a new pair of parameters specific to the central
Meixcan Caribbean. Results suggest there are important regional-scale
variations in allometric growth patterns of lionfish. The observed
differences highlight the importance of using site-specific parameters,
especially when informing invasion management strategies.

@@

The main objective of this work was to compare allometric growth of
lionfish throughout their invasion range. Allometric parameters were
retreived from scientific literature, and an additional pair of
parameters was calculated from filed observations in the central Mexican
Caribbean.

Length-weight relationships (n = 15) identified in literature were
obtained for the North Atlantic (n = 1), Gulf of Mexico (n = 7,), and
Caribbean (n = 8). Information on sampling methods, sex differentiation,
location, and depth ranges of each study was retrieved when available.
Studies were assumed to include both genders if gender was unspecified.
Locations where allometric studies were performed are shown in Figure
\ref{fig:map} and \ref{tab:all_params}.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:map}Locations where allometric growth parameters of
lionfish (\emph{Pterois spp}) have been reported. Circle sizes indicate
sample size from each study, colors indicate the \(b\) coefficient from
Eq. \ref{eq:allometric}.}
\end{figure}

Parameters from the central Mexican Caribbean were obtained from data
collected in 10 sampling sites along the central Mexican Caribbean coast
in 2010 (Table S1). Sampling locations included wall and carpet reefs at
depths between 5.7 m and 38.1 m. All observed lionfish (n = 109) were
collected using hand nets and numbered collection bottles. The use of
hand nets prevented any weight loss due to bleeding and allowed better
representation of small sizes by eliminating gear selectivity. Organisms
were euthanized via pithing and Total Length (TL; mm) and Total Weight
(TW; g) were recorded before freezing organisms.

@@

Where \(a\) is the ponderal index and \(b\) is the scaling exponent or
allometric parameter. When \(b = 3\), it is said that the organism
exhibits a perfect isometric growth. Transforming this equation via
base-10 logarithms:

\begin{equation}
\label{eq:log-alo}
log_{10}(TW) = b\times log_{10}(TL) + log_{10}(a)
\end{equation}

@@

Where \(Y = log_{10}(TW)\), \(m = b\), \(X = log_{10}(TL)\), and
\(c = log_{10}(a)\). Since \(b = m\), we will only use \(b\) throughout
the paper for simplicity. The coefficients (\(c\) and \(b\)) were
estimated with an Ordinary Least Squares Regression and
heteroskedastic-robust standard error correction \citep{zeileis_2004}.
The \(b\) coefficient was tested against the null hypothesis of
isometric growth (\emph{i.e.} \(H_0: b = 3\)). Coefficients were tested
with a two-tailed Student's t, and the significance of the regression
was corroborated with an F-test.

Studies inconsistently defined a as either the ponderal index from Eq.
\ref{eq:allometric} or the y-intercept (\(c\)) from Eq.
\ref{eq:log-alo-trans}. Other studies incorrectly reported parameters as
mm-to-g conversions when they were in fact cm-to-g conversions. Here,
all parameters are reported as TL(mm) to TW(gr) conversions. When
required, values from other studies were transformed for consistency.

Since uncertainty around estimated relationships was not reported in
some of the reviewed studies, it was not possible to test for
statistical differences between relationships. Instead, the 16
length-weight relationships were used to calculate predicted weight from
length observations of the organisms sampled from the central Mexican
Caribbean. Predicted weights were divided by the observed weights to
obtain a ratio. Difference in mean weight ratios across studies were
tested with a one-way Analysis of Variance (ANOVA). All analyses were
performed in R version 3.4.4 \citep{rcore_2018}. Raw data and code used
in this work are available at dryad.org.

@@

The model adjusted to Eq. \ref{eq:log-alo-trans} estimated the
coefficient values at \(b = 3.2347391\) and \(c = -5.4940866\)
(\(R^2 = 0.977\), F(df = 1; 107) = 6928.67, \(p < 0.001\)). The
allometric factor (\(b\)) was significantly different from the value of
isometric growth of \(b = 3\) (\(t(107) = 6.04; p<0.001\)), indicating
that lionfish present allometric growth. More information on model fit
is presented in TableS2. The relationship between TL and TW is presented
in Figure \ref{fig:l-w-carib}.

From this study in the central Mexican Caribbean and the 11
peer-reviewed studies that reported length-weight parameters for
\emph{P. volitans} 15 parameters were identified (Table
\ref{tab:all_params}, Fig \ref{fig:all_allo}). Two studies
\citep{aguilarperera_2016,fogg_2013} reported gender-level and pooled
parameters, while the rest presented pooled results. Reviewed studies
presented information for organisms obtained at depths between 0.5 and
57 m. Three studies explicitly stated that their organisms were sampled
with pole spears \citep{aguilarperera_2016,chin_2016,dahl_2014}, and
five studies mentioned that some of their organisms were obtained with
pole spears (or other type of harpoon) but also hand-held nets or fish
traps
\citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,toledohernndez_2014},
and two studies did not specify how organisms were sampled
\citep{deleon_2013,darling_2011}. \citet{fogg_2013} use spine-less
weight in the length-weight relationship estimation.

@@

There were significant differences in predicted-to-observed weight
ratios estimated for each pair of parameters (F(df = 15; 1728) = 38.26;
p \textless{} 0.001). From all allometric parameters reviewed, those of
\citet{edwards_2014} provided the lowest weight estimates, with an
predicted-to-observed weight ratio of 0.98 \(\pm\) 0.23 (mean \(\pm\)
SD). On the other hand,\citet{barbour_2011} yielded the highest weight
estimates, with a mean (\(\pm\) SD) predicted-to-observed weight ratio
of 1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios and groups
identified by Tukey's HSD (\(\alpha = 0.05\)) are presented in Figure
\ref{fig:bio_ratio}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:l-w-carib}Length-weight relationship for 109
lionfish sampled in the central Mexican Caribbean. Points indicate
samples, solid line indicates curve of best fit, marginal plots
represent the density distribution of each variable.}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-5}\label{tab:all_params}Summary of 16 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter for Eq. 1 (presented in $\times 10^{-5}$), c = y-intercept for Eq. 3, b = exponent or slope for Eq. 1 or Eq. 3, respectively. The $R^2$ column indicates reported model fit.}
\centering
\begin{tabular}[t]{llllrrll}
\toprule
Region & Sex & n & a & b & c & R\textasciicircum{}2 & Reference\\
\midrule
Caribbean & B & 458 & 3.6 & 2.81 & -4.44 & - & Sandel et al., 2015\\
Caribbean & B & 419 & 2.8 & 2.85 & -4.56 & 0.8715 & Chin et al., 2016\\
Caribbean & B & 1450 & 2.3 & 2.89 & -4.64 & 0.96 & de Leon et al., 2013\\
Caribbean & B & 1887 & 0.3 & 3.24 & -5.52 & 0.97 & Edwards et al., 2014\\
Caribbean & B & - & 0.25 & 3.29 & -5.60 & - & Darling et al., 2011\\
\addlinespace
Caribbean & B & 2143 & 0.52 & 3.18 & -5.28 & 0.9907 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 0.8 & 3.11 & -5.10 & 0.958 & Toledo-Hernández et al., 2014\\
Caribbean & B & 109 & 0.32 & 3.23 & -5.49 & 0.9766 & This study\\
GoM & B & 934 & 0.21 & 3.34 & -5.68 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 0.29 & 3.30 & -5.54 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\addlinespace
GoM & F & 67 & 0.12 & 3.47 & -5.93 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & M & 59 & 0.42 & 3.23 & -5.38 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 0.14 & 3.43 & -5.86 & 0.99 & Fogg et al., 2013\\
GoM & M & 119 & 0.27 & 3.31 & -5.57 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 0.68 & 3.14 & -5.17 & 0.94 & Fogg et al., 2013\\
NorthAtlantic & B & 774 & 2.9 & 2.89 & -4.54 & - & Barbour et al.,2011\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 16) for 11
studies and this study. Colors indicate studies from which the
parameters were extracted. Solid lines indicate that the fit was
performed for males and females pooled together. Dotted lines indicate
that the regression was performed on females, and dashed lines indicate
it was performed for males.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:bio_ratio}Violin plot showing the distribution of
predicted to observed weight ratios for 16 pairs of allometric
parameters. Red and blue circles indicate median and mean values,
respectively. Like letters indicate values that do not differ
significantly (Tukey's HSD; p \textless{} 0.05).}
\end{figure}

@@

A new pair of allometric growth parameters for lionfish in the central
Mexican Caribbean are provided. This compliments existing literature for
other sites in the Mexican Caribbean \citep{sabidoitza_2016}.
Additionally, the study identifies regional differences in length-weight
relationships.

The length-weight coefficients estimated in this study were within the
range identified by studies in other regions (Table
\ref{tab:all_params}). However, the ones presented here provide lower
weight estimates for similar lenmgths. Until about TL = 200 mm, there
are no appreciable differences between the parameters for organisms from
the Mexican Caribbean and those for little Cayman \citep{edwards_2014}
and Jamaica \citep{chin_2016}. Yet, for larger organisms (TL
\textgreater{} 270 mm) parameters from Costa Rica \citep{sandel_2015}
and Bonaire \citep{deleon_2013} provide similar estimates to those from
this study. Conversely, these same studies tend to estimate higher
weights --as compared to the ones reported here-- for smaller organisms,
likely due to the lack of small organisms in the samples used to
estimate their parameters.

There are evident differences in weight-at-length between organisms from
the Caribbean and Gulf of Mexico / North-Western Atlantic. Weight
estimates with parameters from the Gulf of Mexico and North-Western
Atlantic tend to be higher than those from the Caribbean. Similar
regional variation has been reported for age-at-length relationships of
this species across the invaded region \citep{fogg_2015}, or when
comparing populations from the invasion and native ranges
\citep{pusack_2016}. These may be driven by genetic differences or
organisms being exposed to distinct environmental conditions. For
example, work on mitochondrial DNA has shown two distinct population
groups, identified as the ``Caribbean group'' and ``Northern Group''
\citep{betancurr_201}. Alternatively, \citet{fogg_2015} suggest that
differences observed in age-at-length ``may bemore related to climate
rather than other biological and ecological factors''. Differences in
weight-at-length could also reflect differences in energy input
(\emph{i.e.} in some regions, lionfish eat more) or differential usage
of this energy (\emph{e.g.} regional differences in predator abundances
lead to different usage of energy), or a combination of both. Future
research should focus on identifying which occurs here.

The results presented in this paper can have major implications for
management. Our results show that, for a given length, using one set of
parameters or the other can result in a threefold increase in estimated
weight. This becomes especially relevant when estimating biomass
available for harvest, predicting effects on local ecosystems, or
evaluating the effectiveness of removal programs. Using site-specific
values provides a more accurate estimate of fish biomass. Future
research should try to use, to the extent possible, parameters
calculated for their region, or use different parameters to provide
upper and lower bounds in their results. At the same time, this
highlights the need for more basic research that furthers our
understanding of lionfish biology. To better manage the invasion, we
must perform research that can describe biologically important
information of lionfish throughout its invasion range
\citep{johnson_2016}.
